---
title: "Data Preprocessing"
author: "David J. Barney"
date: "1/1/2020"
output: html_document
---

# Transformation Glossary
## Single Predictors
* **Centering**: Transform mean to zero
  + Vector of predictor values - average predictor value
* **Scaling**: Transform common standard deviation of one
  + Vector of predictor values / standard deviation of predictor value
* **Skewness**: Asymmetry of distribution on one side of the mean
  + Common to transform with log, square root, or inverse
  + Rule of thumb: ratio of highest:lowest value > 20 indicates significant skewness
  + Box-Cox transformation: monotonic (preserves or reverses order of observations) power transformation. Includes $\lambda$ parameter that specifies the power function: square ($\lambda$ = 2), square root ($\lambda$ = 0.5), inverse ($\lambda$ = -1), and values in between. $\lambda$ is generally estimated using MLE.

## Outliers
* **Spatial sign**: Project predictor values onto multidimensional sphere to minimize effect of outliers
  + Each sample divided by its squared norm
  + Must center and scale all predictors prior to spatial sign transform
  + Predictors transformed as a group, so procedure must be repeated if features are changed

## Data Reduction
* **Principal Component Analysis**: Find linear combinations of predictors that capture greatest possible variance
  + Creates components that are uncorrelated
  + Best practice is to first transform skewed predictors, then center and scale all predictors prior to PCA in order to avoid outsized effect of predictors with greater variation
  
## Removing Predictors
* **Near-zero variance predictors**: Predictors with a single value for the large majority of samples
  + Heuristics for detecting NZVPs:
    + Fraction of unique values over the sample size is low (< 10%)
    + Frequency of modal value:frequency of second most common value is large (~ 20%)
    
* **Multicollinearity**: Significant correlations between multiple predictors
  + Steps to remove minimum number of predictors to ensure all pairwise correlations are below a threshold:
    + Calculate correlation matrix of predictors
    + Identify the two predictors with the largest absolute pairwise correlation (*A* and *B*)
    + Calculate the average correlation between *A* and all other predictors, repeat for *B* and all other predictors
    + If *A* has a larger average correlation, remove it; otherwise, remove *B*
    + Repeat until no absolute correlations are above the threshold
  + Threshold for above procedure depends on your model's sensitivity to between-predictor correlations (e.g. we might establish a threshold of 0.75 for a sensitive model)

# Computing
## Setup
```{r Load packages}
library("AppliedPredictiveModeling")
library("caret")
library("corrplot")
library("e1071")
library("lattice")
```

```{r Prepare initial data}
data("segmentationOriginal") # Load data
segData <- subset(segmentationOriginal, Case == "Train") # Subset to training cases

# Separate cellID, class, and case into vectors
cellID <- segData$Cell
class <- segData$Class
case <- segData$Case

# Remove the cellID, class, and case columns
segData <- segData[, -(1:3)]

# Remove "status" columns
statusColNum <- grep("Status", colnames(segData))
segData <- segData[, -statusColNum]
```

## Transformations
### Examples of individual predictor transforms
#### Examine Skewness
```{r Calculate skewness for predicotrs}
# Apply skewness function over all predictors
skewValues <- apply(segData, 2, skewness)
hist(skewValues)
```

#### Example Box-Cox
```{r Use Box-Cox Transformation }
Ch1AreaTrans <- BoxCoxTrans(segData$AreaCh1)
Ch1AreaTrans
```

```{r Examine original data}
head(segData$AreaCh1)
```

```{r Examine data after Box-Cox transform}
predict(Ch1AreaTrans, head(segData$AreaCh1))
```

```{r Example of a spatial sign transform}
spatialSign(segData[1:3, 1:3])
```

### Principal Components Analysis
```{r Apply centering and scaling to all predictors}
# Create PCA object with centering and scaling
pcaObject <- prcomp(segData, center = TRUE, scale. = TRUE)

# Calculate the cumulative percentage of variance which each component accounts for
percentVariance <- pcaObject$sdev^2/sum(pcaObject$sdev^2)*100
percentVariance[1:3]
```

```{r Examine predictor-component loadings}
head(pcaObject$rotation[, 1:3])
```

### `caret::preProcess` Functionality
```{r Examine outputs from `preProcess`}
# Summarize change from Box-Cox transform, centering, scaling, and PCA
trans <- preProcess(segData, method = c("BoxCox", "center", "scale", "pca"))
trans
```

```{r Create transformed dataframe from `preProcess` outputs using `predict`}
transformed <- predict(trans, segData)
head(transformed[, 1:5])
```

### Filtering
#### Near-Zero Variance Predictors
```{r Filter out near-zero variance predictors with `caret::nearZeroVar`}
nearZeroVar(segData) # No predictors have near-zero variance
```

#### Correlated Predictors
```{r Examine between-predictor correlations}
correlations <- cor(segData)
correlations[1:4, 1:4]
```

```{r Visualize between-predictor correlations}
corrplot(correlations, order = "hclust")
```

```{r Filter correlated predictors}
highCorr <- findCorrelation(correlations, cutoff = 0.75)
length(highCorr)
head(highCorr)
filteredSegData <- segData[, -highCorr]
```

### Dummy Coding
```{r Load `caret::cars` dataset}
data(cars)
type <- c("convertible", "coupe", "hatchback", "sedan", "wagon")
cars$Type <- factor(apply(cars[, 14:18], 1, function(x) type[which(x == 1)]))
carSubset <- cars[sample(1:nrow(cars), 20), c(1, 2, 19)]
carSubset
```


```{r Simple usage of `caret::dummyVars`}
simpleMod <- dummyVars(~Mileage + Type,
                       data = carSubset,
                       levelsOnly = TRUE)
simpleMod
```

```{r Factor expansion using `predict` method}
predict(simpleMod, head(carSubset))
```

```{r Interaction usage of `caret::dummyVars`}
withInteraction <- dummyVars(~Mileage + Type + Mileage:Type,
                             data = carSubset,
                             levelsOnly = TRUE)
withInteraction
```

```{r Factor expansion with interaction using `predict` method}
predict(withInteraction, head(carSubset))
```







